---
title: "StockX Sneaker Data Analysis"
author: "Jarrod Cruz"
date: "2023-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(plotly)
```

## Introduction 
* Today I will be analyzing sneaker resale data from StockX.
* This data set contains a random sample of 99,956 total sales data from 2017-2019.
* It compares Nike Off-White and Yeezy 350 sales in the United States.

### Problems/Questions To Be Answered
* What shoes are the most popular?
* Which shoes are the most/least profitable?
* Does region or size affect the profit?

## Reading the Data
### Data Wrangling and Cleaning
```{r, }
library(readr)
sneaker_data <- read_csv("StockX-Data-Contest-2019-3.csv",  )
head(sneaker_data)

```
From this we can see that there are a few formatting issues. The column types are wrong, as the price variables and date variables need to be numeric. Additionally, I will change the names of the columns for workability. 



```{r}

colnames(sneaker_data) <- c('Order_Date',
                            'Brand',
                            'Shoe_Name',
                            'Resale_Price',
                            'Retail_Price',
                            'Release_Date',
                            'Shoe_Size',
                            'Buy_Region')

sneaker_data2 <- sneaker_data %>%
  mutate(Order_Date = as.Date(Order_Date, format = "%m/%d/%Y")) %>%
  mutate(Resale_Price = parse_number(Resale_Price)) %>%
  mutate(Retail_Price = parse_number(Retail_Price)) %>%
  mutate(Release_Date = as.Date(Release_Date, format = "%m/%d/%Y")) %>%
  mutate(Shoe_Size = as.numeric(Shoe_Size))
head(sneaker_data2)
sum(is.na(sneaker_data2))

```
The column types are all correct, and there are no missing values in the data set.
Now that the data is properly manipulated, we can begin exploring it.

## Data Exploration

### Average Resale Price of Each Shoe By Brand

```{r, echo=FALSE}
avgResaleBySneaker<- sneaker_data2 %>%
  select(
  Brand, 
  Shoe_Name, 
  Resale_Price,
  Retail_Price,
  Shoe_Size,
  Shoe_Name) %>%
  group_by(Shoe_Name) %>%
  summarize(Brand, Retail_Price, Average_Resale_Price = mean(Resale_Price))

NikeSneakers <- avgResaleBySneaker %>%
  filter(Brand=="Off-White") %>%
  arrange(-Average_Resale_Price)
   

YeezySneakers <- avgResaleBySneaker %>%
  filter(Brand=="Yeezy") %>%
  arrange(-Average_Resale_Price) 
  
 
#ggplot(avgResaleBySneaker,aes(Brand,Average_Resale_Price))+geom_point()
x = avgResaleBySneaker$Brand
y = avgResaleBySneaker$Average_Resale_Price
xax <- list(
  title = "Brand",
  titlefont = list(family="Modern Computer Roman")
)
yax <- list(
  title = "Average Profit",
  titlefont = list(family="Modern Computer Roman")
)

figure <- plot_ly(x=x, y=y, type="scatter", mode="markers",name="markers") %>%
   
   layout(xaxis = xax, yaxis = yax)
figure
```

It is interesting to see how much aftermarket value some sneakers have. Generally speaking, most of the Nike Off White shoes resale for double their retail value.

### Most Popular Sneaker on Average by Brand

The most popular sneaker in this case will be the one with the highest resale value. We will split the shoes by brand, to see which has the more sought-after shoe.

```{r }

mostPopularNikeSneaker <- head(NikeSneakers,1) %>% print()
```

The most popular Nike Off-White sneaker is the Air Jordan 1 Retro High Off White in the White color.

```{r}
mostPopularYeezySneaker <- head(YeezySneakers,1) %>% print()

```


The Most popular Yeezy sneaker is the Adidas Yeezy Boost 350 Low in the TurtleDove color.

### Most Profitable Shoe

```{r }
profitOfShoe <- avgResaleBySneaker %>% summarize(Brand, Shoe_Name, Retail_Price, Average_Resale_Price, Average_Profit = Average_Resale_Price - Retail_Price) %>% unique() 
head(profitOfShoe %>% arrange(-Average_Profit),1)


```

The Nike Off-White Air Jordan 1 in the white colorway has the highest profit of $1636.

### Least Profitable Shoe 

```{r}
head(profitOfShoe %>% arrange(+Average_Profit),1)
```
The Adidas Yeezy Boost 250 V2 in the Sesame colorway has the lowest profit of $44.

## Yeezy Sneaker Average Profit and Retail Interactive Bar Graph
Hover over a color to see the resale value, name, and retail cost of an Adidas Yeezy Sneaker.
```{r, echo=FALSE}
a <- ggplot(profitOfShoe %>%
filter(Brand=="Yeezy") %>%
arrange(-Average_Profit) , aes(x=Retail_Price,y=Average_Profit)) + 
geom_bar(aes(fill = Shoe_Name), stat='identity', width = 5) + labs(title = "Interactive Graph of Yeezy Profit",   x = "Retail Price", y = "Cumulative Profit") +
  theme(
        axis.text = element_text(size = 6),
        plot.title = element_text(size = 12),
        legend.position = "right",
        legend.direction = "vertical",
        legend.text = element_text(size = 6),
        legend.key.size = unit(0.1, "cm"))

ggplotly(a)


```

The y-axis displays the cumulative profit for all sneakers in the category, but hovering over the color of a sneaker on the graph will represent the average profit of that shoe.

## Interactive Plot of Nike Off-White Sneaker Profit

Hover over a color to see the resale value, name, and retail cost of a Nike Off-White sneaker. 
```{r, echo=FALSE}
b <- ggplot(profitOfShoe %>%
filter(Brand=="Off-White") %>%
arrange(-Average_Profit) , aes(x=Retail_Price,y=Average_Profit)) + 
geom_bar(aes(fill = Shoe_Name), stat='identity', width = 5) + labs(title = "Interactive Graph of Off-White Profit",   x = "Retail Price", y = "Cumulative Profit") +
  theme(
        axis.text = element_text(size = 6),
        plot.title = element_text(size = 12),
        legend.position = "right",
        legend.direction = "vertical",
        legend.text = element_text(size = 6),
        legend.key.size = unit(0.1, "cm"))

ggplotly(b)


```

The y-axis displays the cumulative profit for all sneakers in the category, but hovering over the color of a sneaker on the graph will represent the average profit of that shoe.


## What Factors Can Affect Profit?

Now that we have discovered the most and least profitable shoes, it seems fitting to explore and determine what role certain factors may or may not have on profitability.

### Shoe Size?

Let's see if there is a relationship between shoe size and profit by running a simple linear regression on shoe size and average profit.

#### Visualization of Average Profit and Shoe Size

```{r, echo=FALSE}
sneakerSizePlotData <- sneaker_data2 %>%
  select(
  Brand, 
  Shoe_Name, 
  Resale_Price,
  Retail_Price,
  Shoe_Size,
  Buy_Region,
  Shoe_Name) %>%
  group_by(Shoe_Size) %>%
  summarize(Shoe_Size, Retail_Price, Resale_Price,Avg_Retail = mean(Retail_Price), Avg_Resale = mean(Resale_Price), Profit = Resale_Price - Retail_Price, Avg_Profit = Avg_Resale - Avg_Retail)

#profit_by_Size <-
sizePlot =ggplot(sneakerSizePlotData, aes(x = Shoe_Size, y= Avg_Profit)) +  geom_point() + stat_smooth()

sizePlot
```

From this graph we can see that there is a small trend at the end for larger shoe sizes and profit.

#### Simple Linear Regression of Shoe Size and Average Profit




```{r}

sizeModel <- lm(Avg_Profit ~ Shoe_Size, data = sneakerSizePlotData)
sizeModel
ggplot(sneakerSizePlotData, aes(x = Shoe_Size, y= Avg_Profit)) +  geom_point() + stat_smooth(method = lm)

```

Displayed above is the model that was computed and the graph with a regression line. We can now summarize and determine if there is a relationship.

```{r}
summary(sizeModel)
```
Based on this, there does not seem to be significant difference to suggest that the model has any predictive ability. There does not seem to be a relationship between shoe size and profit.

### Region?

Now that we have examined the effect of shoe size on average profit, let's take a look at buyer region. Following a similar procedure,
we can see if region affects the profit of a shoe by running a simple linear regression.

#### Visualization of Average Profit by Buyer Region

```{r,echo=FALSE}
sneakerRegionPlotData <- sneaker_data2 %>%
  select(
  Brand, 
  Shoe_Name, 
  Resale_Price,
  Retail_Price,
  Shoe_Size,
  Buy_Region,
  Shoe_Name) %>%
  group_by(Buy_Region) %>%
  summarize(Buy_Region, Retail_Price, Resale_Price,Avg_Retail = mean(Retail_Price), Avg_Resale = mean(Resale_Price), Profit = Resale_Price - Retail_Price, Avg_Profit = Avg_Resale - Avg_Retail)

#profit_by_Size <-
regionPlot =ggplot(sneakerRegionPlotData, aes(x = Buy_Region, y= Avg_Profit)) +  geom_point() + stat_smooth()

regionPlot + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

There already does not seem to be a trend with the data.

#### Simple Linear Regression of Average Profit and Buyer Region
```{r}
regionModel <- lm(Avg_Profit ~ Buy_Region, data = sneakerRegionPlotData)
regionModel
ggplot(sneakerRegionPlotData, aes(x = Buy_Region, y= Avg_Profit)) +  geom_point() + stat_smooth(method = lm) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



The program is unable to create a regression line. We can still analyze the summary of the model.
```{r}
summary(regionModel)

```


There does not seem to be a relationship between the two variables.

## Conclusions

* The Nike Off-White sneakers are generally more profitable than the Adidas Yeezy 350 v2.
* Shoe size and buyer region did not correlate with increased profits.
* On average, Delaware made the highest profits on sneaker sales.
